Screening of airborne radioxenon measurements for Nuclear-Test-Ban Treaty verification

### 2022-09-23 paper/paper.tex
Added preliminary results section to paper.

\documentclass{article}
\usepackage{amsmath, amssymb, amsthm, dsfont, physics, hyperref, cleveref}
\usepackage[margin=1.5cm]{geometry}
\usepackage{biblatex}

@@

>>>>

\documentclass{article}
\usepackage{amsmath, amssymb, amsthm, dsfont, physics, hyperref, cleveref}
\usepackage{graphicx}
\usepackage[margin=1.5cm]{geometry}
\usepackage{biblatex}

@@

Before applying any optimization procedure, we must establish a baseline. This is done by filling in the daily source emissions $x$ in \cref{eq: initial model} with their yearly averages. The simulated observations in \cref{fig: y-histogram baseline} seem to be biased towards lower values than expected given the actual measurements.\\

\begin{figure}
	\centering
	\includegraphics[width=0.5\linewidth]{../plots/original_simulated_observation.png}
	\caption{Distribution of the simulated observation using the yearly averaged emissions at all times.}
	\label{fig: y-histogram baseline}
\end{figure}

In order to quantify the performance of the simulations, we define some metrics. First of all, the fraction of the simulated observations which lie within a factor $2$ of the actual observations. In the ideal case, all simulated observations $\hat{y}$ would be exactly the same as the observed ones $y$. This metric allows for some small scaling factor as error.
\begin{equation}
	\text{FAC2}:=P(0.5\leq\hat{y}/y\leq 2)
\end{equation}

The previous metric only takes into account the simulated observations within a small area. Therefore it cannot tell us whether some general bias is present. To investigate this, we use the geometric mean bias. The more biased the ratio is, the further this value will lie from $1$. This is because it defines some mean ratio of actual observations and simulated ones.
%Nitpick Thomas: err, I did not actually check the formulae for the metrics closely during the week. It seems slightly weird taking the exponential, as we mostly care about the scalings (in log space). It might complicate the explanation: "needs to be close to 1 (in some logarithmic way)". Also, why do we have this minus sign in front?
\begin{equation}
	\text{GM}:=e^{-E(ln(\hat{y}/y))}
\end{equation}

\begin{figure}
	\centering
	\begin{minipage}{.48\textwidth}
		\centering
		\includegraphics[width=0.9\linewidth]{../plots/hfun_weights_lambda10.png}
		\caption{Histogram of the optimized weights $W$, using a large regularization parameter $\lambda=10$.}
		\label{fig: optimized weights lambda=10}
	\end{minipage}\hfill
	\begin{minipage}{.48\textwidth}
		\centering
		\includegraphics[width=0.9\linewidth]{../plots/hfun_simulated_observations_lambda10.png}
		\caption{Histogram of the ratio of simulated observations $\hat{y}/y$, using a large regularization parameter $\lambda=10$.}
		\label{fig: simulated observations lambda=10}
	\end{minipage}
\end{figure}

\begin{figure}
	\centering
	\begin{minipage}{.48\textwidth}
		\centering
		\includegraphics[width=0.9\linewidth]{../plots/hfun_weights_lambda1.png}
		\caption{Histogram of the optimized weights $W$, using a medium regularization parameter $\lambda=1$.}
		\label{fig: optimized weights lambda=1}
	\end{minipage}\hfill
	\begin{minipage}{.48\textwidth}
		\centering
		\includegraphics[width=0.9\linewidth]{../plots/hfun_simulated_observations_lambda1.png}
		\caption{Histogram of the ratio of simulated observations $\hat{y}/y$, using a medium regularization parameter $\lambda=1$.}
		\label{fig: simulated observations lambda=1}
	\end{minipage}
\end{figure}

\begin{figure}
	\centering
	\begin{minipage}{.48\textwidth}
		\centering
		\includegraphics[width=0.9\linewidth]{../plots/hfun_weights_lambda01.png}
		\caption{Histogram of the optimized weights $W$, using a small regularization parameter $\lambda=0.1$.}
		\label{fig: optimized weights lambda=0.1}
	\end{minipage}\hfill
	\begin{minipage}{.48\textwidth}
		\centering
		\includegraphics[width=0.9\linewidth]{../plots/hfun_simulated_observations_lambda01.png}
		\caption{Histogram of the ratio of simulated observations $\hat{y}/y$, using a small regularization parameter $\lambda=0.1$.}
		\label{fig: simulated observations lambda=0.1}
	\end{minipage}
\end{figure}

Starting with the results when using a high regularization parameter $\lambda=10$, we observe that the weights are nearly $1$ everywhere (see \cref{fig: optimized weights lambda=10}). This makes sense, as the cost function in this case heavily penalizes the weights for deviating from $1$. As a result, the histogram of simulated results (see \cref{fig: simulated observations lambda=10}) is nearly identical to the baseline. Lowering the regularization parameter $\lambda=1$, we find that the weights have more freedom to deviate from $1$ (see \cref{fig: optimized weights lambda=1}) and the bias on the ratio $\hat{y}/y$ decreases (see \cref{fig: simulated observations lambda=1}). This is a direct result of lowering the influence of the regularization term. Finally, lowering the regularization parameter $\lambda=0.1$ even further, the weights get even more freedom (see \cref{fig: optimized weights lambda=0.1}), while the ratio $\hat{y}/y$ finally seems to be centered around $1$. In \cref{table: evaluated metrics}, a summary of the metrics is found, evaluated at the different values for the regularization parameter. Summarized, we see the most major change in the geometric mean bias GM as the regularization parameter decreases. This might lead us to believe that the best result is obtained by choosing $\lambda$ as small as possible. However, a small value for $\lambda$ results in larger deviations from the mean emission; these can become greater than can what be expected in real life TODO REF PAPER typical variation. Thus a fine balance between the prediction error and the regularization condition is still needed.\\

\begin{table}
	\centering
	\caption{Metrics applied to the simulated observations in case of different regularization parameters $\lambda$.}\label{table: evaluated metrics}
\begin{tabular}{|l||c|c|}
	\hline
	&FAC2&GM\\\hline
	baseline&$31.7\%$&3.050\\\hline
	$\lambda=10$&$32.2\%$&2.984\\\hline	
	$\lambda=1$&$40.5\%$&1.776\\\hline	
	$\lambda=0.1$&$40.8\%$&1.216\\\hline
\end{tabular}
\end{table}
